---
title: "Genetic clustering for nutrition traits info"
output: html_notebook
---

. 
Accessing genomic data and nutrition data

```{r, message=F, warning=F}
accessions2 <- read.csv("~/Desktop/PI4 2017/myproject/Team5/data/df_traits_nutrition.csv")
genesys_data2 <- accessions2 %>%
  mutate(acceNumb = gsub("PI ", "PI", acceNumb))
```

```{r}
genesys_data_sorted2 <- genesys_data2[order(genesys_data2$acceNumb),]

reference = genome_table@fix
View(reference)
gt_file = genome_table@gt
meta_file = genome_table@meta

colnames(gt_file)<-gsub(".MERGE", "",colnames(gt_file))

genesys_data_sorted2 <- genesys_data2[order(genesys_data2$acceNumb),]
```

```{r}
match_vcf2<-colnames(gt_file) %in% genesys_data_sorted2$acceNumb
colnames(gt_file)[match_vcf2]



match_genesys2<- genesys_data_sorted2$acceNumb  %in% colnames(gt_file)
genesys_data_sorted2[match_genesys2,]$acceNumb

genesys_data_match2 <- genesys_data_sorted2[match_genesys2,]
vcf_gt_match2 <- gt_file[,match_vcf2]


```


```{r}
d_matrix <- matrix(,nrow = 48,ncol=48)
for(i in 1:48){
  for(j in i:48){
    d_matrix[i,j] <- cityblock(vcf_gt_match2[,i],vcf_gt_match2[,j])
    d_matrix[j,i] <-d_matrix[i,j]
  }
}
heatmap(d_matrix, sym=TRUE)
ct_tree<-hclust(as.dist(d_matrix))

plot(ct_tree,main='Nutrition trait genomic clusters')

```
CLustering for euclidean distance
```{r}
ed_matrix <- matrix(,nrow = 48,ncol=48)
for(i in 1:48){
  for(j in i:48){
    ed_matrix[i,j] <- euclidean(vcf_gt_match2[,i],vcf_gt_match2[,j])
    ed_matrix[j,i] <-ed_matrix[i,j]
  }
}
heatmap(ed_matrix, sym=TRUE)
ct_tree <-hclust(dist(ed_matrix,method = "euclidean")) %>% 
  as.dendrogram() %>% 
  set("labels_cex", 0.6) %>%
  set("labels_col",value = 2:4,k=6) %>%
  set("branches_k_color", value = 2:4, k = 6) %>%
  set("labels",colnames(vcf_gt_match2)) %>%
  plot(main='Nutrition trait genomic clusters') 

png("myproject/Team5/48_trait_euclidean_genomic_clusters.png")
ct_tree <-hclust(dist(ed_matrix,method = "euclidean")) %>% 
  as.dendrogram() %>% 
  set("labels_cex", 0.6) %>%
  set("labels_col",value = 2:4,k=6) %>%
  set("branches_k_color", value = 2:4, k = 6) %>%
  set("labels",colnames(vcf_gt_match2)) %>%
  plot(main='Nutrition trait genomic clusters') 
dev.off()  
```
clustering for hamming distance

```{r}
hd_matrix48 <- matrix(,nrow = 48,ncol=48)
for(i in 1:48){
  for(j in i:48){
    hd_matrix48[i,j] <- hamming(vcf_gt_match2[,i],vcf_gt_match2[,j])
    hd_matrix48[j,i] <-hd_matrix48[i,j]
  }
}
heatmap(hd_matrix, sym=TRUE)
ct_tree48 <-hclust(dist(hd_matrix48,method = "manhattan")) %>% 
  as.dendrogram() %>% 
  set("labels_cex", 0.6) %>%
  set("labels_col",value = 2:4,k=6) %>%
  set("branches_k_color", value = 2:4, k = 6) %>%
  set("labels",colnames(vcf_gt_match2)) %>%
  plot(main='Nutrition trait genomic clusters') 

png("myproject/Team5/48_trait_binary_genomic_clusters.png")
ct_tree48_2 <-hclust(dist(hd_matrix,method = "manhattan")) %>% 
  as.dendrogram() %>% 
  set("labels_cex", 0.6) %>%
  set("labels_col",value = 2:4,k=6) %>%
  set("branches_k_color", value = 2:4, k = 6) %>%
  set("labels",colnames(vcf_gt_match2)) %>%
  plot(main='Nutrition trait genomic clusters') 
dev.off()  
```

clustering for manhattan distance

```{r}
cd_matrix48 <- matrix(,nrow = 48,ncol=48)
for(i in 1:48){
  for(j in i:48){
    cd_Matrix48[i,j] <- cityblock(vcf_gt_match2[,i],vcf_gt_match2[,j])
    cd_matrix48[j,i] <-cd_matrix48[i,j]
  }
}
heatmap(hd_matrix, sym=TRUE)
ct_tree <-hclust(dist(cd_matrix,method = "manhattan")) %>% 
  as.dendrogram() %>% 
  set("labels_cex", 0.6) %>%
  set("labels_col",value = 2:4,k=6) %>%
  set("branches_k_color", value = 2:4, k = 6) %>%
  set("labels",colnames(vcf_gt_match2)) %>%
  plot(main='Nutrition trait genomic clusters') 

png("myproject/Team5/48_trait_manhattan_genomic_clusters.png")
ct_tree <-hclust(dist(cd_matrix,method = "manhattan")) %>% 
  as.dendrogram() %>% 
  set("labels_cex", 0.6) %>%
  set("labels_col",value = 2:4,k=6) %>%
  set("branches_k_color", value = 2:4, k = 6) %>%
  set("labels",colnames(vcf_gt_match2)) %>%
  plot(main='Nutrition trait genomic clusters') 
dev.off()  
```

```{bash}
pwd
```

```{r}

files <- dir( pattern = 'csv', full.names = TRUE)
accessions <- read.csv(files[6])
files <- files[c(1:5,7:36)]
tables <- gsub(".csv", "", basename(files))

library(dplyr)
for(i in seq_along(files)){
  next_table <- read.csv(files[i])
  next_table$datasetId <- NULL
  colnames(next_table)[colnames(next_table) == 'value'] <- tables[i]
  accessions <- left_join(accessions, next_table, by = 'genesysId')
}

genesys_data <- accessions %>%
  mutate(acceNumb = gsub("PI ", "PI", acceNumb))
```

```{r}
genesys_data_sorted <- genesys_data[order(genesys_data$acceNumb),]

reference = genome_table@fix
View(reference)
gt_file = genome_table@gt
meta_file = genome_table@meta

colnames(gt_file)<-gsub(".MERGE", "",colnames(gt_file))

genesys_data_sorted <- genesys_data[order(genesys_data$acceNumb),]
```

```{r}
match_vcf<-colnames(gt_file) %in% genesys_data_sorted$acceNumb
colnames(gt_file)[match_vcf]



match_genesys<- genesys_data_sorted$acceNumb  %in% colnames(gt_file)
genesys_data_sorted[match_genesys,]$acceNumb

genesys_data_match <- genesys_data_sorted[match_genesys,]
vcf_gt_match <- gt_file[,match_vcf]


```


manhattan 
```{r}
d_matrix2 <- matrix(,nrow = 19,ncol=19)
for(i in 1:19){
  for(j in i:19){
    d_matrix2[i,j] <- cityblock(vcf_gt_match[,i],vcf_gt_match[,j])
    d_matrix2[j,i] <-d_matrix2[i,j]
  }
}

ct_tree <-hclust(dist(d_matrix,method = "manhattan")) %>% 
  as.dendrogram() %>% 
  set("labels_cex", 0.6) %>%
  set("branches_k_color", value = 2:4, k = 3) %>%
  set("labels",colnames(vcf_gt_match)) %>%
  plot(main='Nutrition trait genomic clusters') 

png("myproject/Team5/19_trait_manhattan_genomic_clusters.png")
ct_tree <-hclust(dist(d_matrix,method = "manhattan")) %>% 
  as.dendrogram() %>% 
  set("labels_cex", 0.6) %>%
  set("branches_k_color", value = 2:4, k = 3) %>%
  set("labels",colnames(vcf_gt_match)) %>%
  plot(main='Nutrition trait genomic clusters') 
dev.off()  

```

euclidean
```{r}
d_matrix <- matrix(,nrow = 19,ncol=19)
for(i in 1:19){
  for(j in i:19){
    d_matrix[i,j] <- hamming(vcf_gt_match[,i],vcf_gt_match[,j])
    d_matrix[j,i] <-d_matrix[i,j]
  }
}

d_matrix2 <- matrix(,nrow = 19,ncol=19)
for(i in 1:19){
  for(j in i:19){
    d_matrix2[i,j] <- cityblock(vcf_gt_match[,i],vcf_gt_match[,j])
    d_matrix2[j,i] <-d_matrix2[i,j]
  }
}


ct_tree3 <-hclust(dist(d_matrix,method = "manhattan")) %>% 
  as.dendrogram() %>% 
  set("labels_cex", 0.6) %>%
  set("branches_k_color", value = 2:4, k = 3) %>%
  set("labels",colnames(vcf_gt_match)) %>%
  set("labels_col",value = 2:4,k=3) 
  #plot(main='Nutrition trait genomic clusters') 

#png("myproject/Team5/19_trait_euclidean_genomic_clusters.png")
ct_tree4 <-hclust(dist(d_matrix2,method = "manhattan")) %>% 
  as.dendrogram() %>% 
  set("labels_cex", 0.6) %>%
  set("branches_k_color", value = 2:4, k = 3) %>%
  set("labels",colnames(vcf_gt_match)) %>%
  set("labels_col",labels_col(ct_tree3)[order(labels(ct_tree4))]) 
  #plot(main='Nutrition trait genomic clusters') 
#dev.off()  
 
png("dual48clusters.png")
par(mfrow=c(1,2))
plot(ct_tree3,main="nutrition trait hamming clusters",horiz=TRUE)
plot_horiz.dendrogram(ct_tree4,main="nutrition trait manhattan clusters",side=TRUE)
dev.off()

```

distance vector
```{r}

d_matrix <- matrix(,nrow = 19,ncol=19)
for(i in 1:19){
  for(j in i:19){
    d_matrix[i,j] <- cityblock(vcf_gt_match[,i],vcf_gt_match[,j])
    d_matrix[j,i] <-d_matrix[i,j]
  }
}
d_vector <- vector()
for(i in 1:19){
    d_vector[i] <- cityblock(vcf_gt_match[,i],rep('0|0',70693))
}

ct_tree <-hclust(dist(d_matrix,method = "manhattan"))
```

```{r}
genesys_data_match[,40] <- d_vector
```

```{r}
genesys_matrix <- data.matrix(genesys_data_match[,c('anthesis1','exsertion','plantcol','planthgt')], rownames.force = NA)
rownames(genesys_matrix) <- genesys_data_match[,'V40']
heatmap(scale(genesys_matrix),Rowv = as.dendrogram(ct_tree))
```
Other traits were not in numerical form, but they seemed varied and interesting enough to quantify in some way.  These are, endosprmco, endosprmtx,glumecolor,heightunif,kercover,kershape,nodaltillr,plantcol,rust.  The following is a quantification of the qualifiable data: 

```{r}
qual_data_match<- genesys_data_match[, c("glumecolor","heightunif","kercover","kershape","nodaltillr","plantcol","rust")]
qual_data_match
opts <- list( c("glumecolor","Straw","Tan","Red","Purple"),
              c("heightunif"," (2.0) (1.0 = Very uniform, 5.0 = Not uniform)",
                " (1.5) (1.0 = Very uniform, 5.0 = Not uniform)",
                "Very uniform"),
              c("kercover","25% kernel covered","50% kernel covered","75% kernel covered"),
              c("kershape","Oval","Round","Variable"),
              c("nodaltillr","No","Yes"),
              c("plantcol","Mixed","Purple ","Purple-red","Red ","Red-purple","Tan"),
              c("rust","(1.5) (1.0 = Resistant, 5.0 = Susceptible)",
                "(2.0) (1.0 = Resistant, 5.0 = Susceptible)",
                "(2.5) (1.0 = Resistant, 5.0 = Susceptible)",
                "(3.0) (1.0 = Resistant, 5.0 = Susceptible)",
                "(3.5) (1.0 = Resistant, 5.0 = Susceptible)",
                "(4.0) (1.0 = Resistant, 5.0 = Susceptible)"))

opts_sublist<- list(c(0,.25,.5,.75,1), c(0,0,.5,1),c(0,0,.5,1),c(0,0,.5,1),
                c(0,0,1),c(0,0,1,.8,.4,.6,.2),c(0,0,.2,.4,.6,.8,1))
  

              
for(i in 1:length(opts)){
  
  for ( j in 2:length(opts[[i]]) ){
    
    qual_data_match[,opts[[i]][1]]<- gsub(opts[[i]][j],opts_sublist[[i]][j],qual_data_match[,opts[[i]][1]],fixed=TRUE)
 }
}

qual_data_match
```

```{r}
genesys_data_match[, c("glumecolor","heightunif","kercover","kershape","nodaltillr","plantcol","rust")]<-qual_data_match

genesys_matrix <- data.matrix(genesys_data_match[,c('anthesis1','exsertion',                                            "glumecolor","heightunif","kercover","kershape","nodaltillr",'plantcol','planthgt','rust')], rownames.force = NA)

qual_matrix<-data.matrix(genesys_data_match[, c("glumecolor","heightunif","kercover","kershape","nodaltillr","plantcol","rust")], rownames.force = NA)
rownames(genesys_matrix) <- genesys_data_match[,'V40']
heatmap(qual_matrix,Rowv = as.dendrogram(ct_tree))

```

